---
title: "Random Forest - Simple example"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Some links about random forest variable importance metrics

 - [useful blog post](https://www.r-bloggers.com/be-aware-of-bias-in-rf-variable-importance-metrics/)
   - suggests permutation importance as more robust
   - this accounts for highly correlated predictor variables, but is super slow
 - [useful slides by authors of party package](https://www.statistik.uni-dortmund.de/useR-2008/slides/Strobl+Zeileis.pdf) 

## Load packages

```{r}
library(tidyverse)
library(tidymodels)
library(ranger)
```

## Load data

```{r}
df <- read_csv(file.path("..", "results", "00_SelectGagesForAnalysis_GageSampleAnnual.csv"))

names(df)
```

## build a simple model

We will try to predict `annualnoflowdays` based on some climate and land use variables.

```{r}
# define predictors
predictors <- c("p_mm_cy", "pet_mm_cy", "T_max_c_cy", "p.pet_cy", 
                "lulc_water_prc", "lulc_dev_prc", "lulc_wetland_prc",
                "lulc_forest_prc", "lulc_barren_prc", "lulc_grass_prc", "lulc_ag_prc")

# make data frame with some descriptive variables, the target and the predictor variables
fit_data_in <- 
  df %>% 
  dplyr::select(c("gage_ID", "currentclimyear", "annualnoflowdays", all_of(predictors))) %>% 
  subset(complete.cases(.)) # remove any row with an NA value

# set up the model engine
rf_engine <- 
  rand_forest() %>% 
  set_engine("ranger", 
             num.threads = (parallel::detectCores() - 1),
             importance = "permutation") %>% 
  set_mode("regression")

# set up some preprocessing steps
rf_recipe <-
  fit_data_in %>% 
  # predict annualnoflowdays based on all the other variables
  recipe(annualnoflowdays ~ .) %>%
  # tell model which variables to not use in the prediction (ID variables)
  update_role(gage_ID, currentclimyear, new_role = "ID")

# combine into a workflow
rf_workflow <-
  workflow() %>% 
  add_model(rf_engine) %>% 
  add_recipe(rf_recipe)
    
# fit model
rf_fit <- 
  rf_workflow %>% 
  fit(data = fit_data_in)

```

## look at model performance

```{r}
# extract predictions from model
fit_data_in$predicted <- predict(rf_fit, fit_data_in)$.pred

# plot
ggplot(fit_data_in, aes(x = predicted, y = annualnoflowdays, color = currentclimyear)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  labs(title = "Simple model for annual no flow days") +
  scale_x_continuous(name = "Predicted", expand = c(0,0)) +
  scale_y_continuous(name = "Observed", expand = c(0,0)) +
  scale_color_viridis_c(name = "Year") +
  coord_equal() +
  theme_bw()

```

## investigate variable importance

```{r}
# extract  importance from model
var_imp <- tibble(predictor = names(pull_workflow_fit(rf_fit)$fit$variable.importance),
                  IncMSE = pull_workflow_fit(rf_fit)$fit$variable.importance) %>% 
  arrange(-IncMSE) %>%  # order most to least importance (higher value = greater influence on predictors)
  mutate(predictor_factor = factor(predictor, levels = .$predictor)) # make a factor so they plot in order

ggplot(var_imp, aes(x = predictor_factor, y = IncMSE)) +
  geom_col() +
  theme_bw() +
  theme(axis.text = element_text(angle = 90, hjust = 1)) +
  NULL


```